### 2025-04-11 multiple-regression.qmd
20250410 - multiple regression

```{r}
library("petersenlab")
library("rms")
library("car")
library("lavaan")
library("mice")
library("brms")
library("tidyverse")
library("knitr")
```

@@

@@

The histogram for target share is in @fig-histogramFantasyPoints

@@

@@

@@

```{r}
linearRegressionModel_outcomeTransformed <- lm(
  fantasyPoints ~ age + height + weight + I(log(target_share + 1)),
  data = player_stats_seasonal %>% filter(position %in% c("WR")),
  na.action = "na.exclude"
)

@@

@@

@@

@@

Then, what are the regression coefficients?
We do not know what are the correct regression coefficients because each of the possibilities fits the data equally well.
Thus, when estimating the regression model, we could obtain arbitrary estimates of the regression coefficients with an enormous standard error around each estimate.
In general, multicollinearity increases the uncertainty (i.e., standard errors and confidence intervals) around the parameter estimates.
Any [predictor variables](#sec-correlationalStudy) that have a correlation above ~ $r = .30$ with each other could have an impact on the confidence interval of the regression coefficient.
As the correlations among the [predictor variables](#sec-correlationalStudy) increase, the chance of getting an arbitrary answer increases, sometimes called "bouncing betas."
So, it is important to examine a correlation matrix of the [predictor variables](#sec-correlationalStudy) before putting them in the same regression model.
You can also examine indices such as variance inflation factor (VIF).

@@

@@

@@

@@

rms::robcov(rms::ols(
  fantasyPoints ~ height,
  data = player_stats_seasonal_subset,
  x = TRUE,
  y = TRUE),
  cluster = player_stats_seasonal_subset$player_id) #account for nested data within player
```

@@

>>>>

```{r}
library("petersenlab")
library("rms")
library("car")
library("caret")
library("lme4")
library("performance")
library("lavaan")
library("mice")
library("miceadds")
library("brms")
library("tidyverse")
library("knitr")
```

@@

To evaluate the shape of the association between the predictor variables and the outcome variable, we can examine scatterplots, residual plots, marginal model plots, and added-variable plots.

To evaluate homoscedasticity, we can evaluate residual plots (@fig-residualPlots).
In residual plots, homoscedasticity is identified as a constant spread of residuals versus fitted values—the residuals do not show a fan or cone shape.

To determine if residuals are correlated by a grouping level, we can examine the proportion of variance that is attributable to the grouping level using the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).
The greater the ICC value, the more variance is accounted for by the grouping level, and the more the residuals are intercorrelated.

We can examine whether residuals are normally distributed using quantile–quantile (QQ) plots and probability–probability (PP) plots.
If the residuals are normally distributed, they should stay close to the diagonal reference line.

@@

The histogram for raw fantasy points is in @fig-histogramFantasyPoints.

@@

Let's use a Yeo-Johnson transformation of fantasy points:

```{r}
yjTransformed <- caret::preProcess(
  player_stats_seasonal["fantasyPoints"],
  method = c("YeoJohnson"))

yjTransformed

player_stats_seasonal$fantasyPoints_transformed <- predict(
  yjTransformed,
  newdata = player_stats_seasonal["fantasyPoints"])$fantasyPoints
```

@@

hist(player_stats_seasonal$fantasyPoints_transformed[which(player_stats_seasonal$position == "WR")])
```

@@

```{r}
linearRegressionModel_outcomeTransformed <- lm(
  fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
  data = player_stats_seasonal %>% filter(position %in% c("WR")),
  na.action = "na.exclude"
)

@@

The residuals show more constant variance after transforming the outcome variable.

@@

To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).

```{r}
mixedModel <- lmer(
  fantasyPoints_transformed ~ 1 + (1 | player_id),
  data = player_stats_seasonal)

performance::icc(mixedModel)
```

The ICC indicates that over half of the variance is attribute to between-player variance, so it would be important to account for the player-specific variance using a [mixed model](#sec-mixedModels).
For simplicity, we focus on multiple regression models in this chapter; mixed models are described in @sec-mixedModels.

@@

We can examine whether residuals are normally distributed using quantile–quantile (QQ) plots and probability–probability (PP) plots, as in Figures [-@fig-qqPlot] and [-@fig-ppPlot].
If the residuals are normally distributed, they should stay close to the diagonal reference line.

```{r}
#| label: fig-qqPlot
#| fig-cap: "Residual Plots After Transformation of Fantasy Points."
#| fig-alt: "Residual Plots After Transformation of Fantasy Points."

car::qqPlot(
  linearRegressionModel_outcomeTransformed,
  main = "QQ Plot",
  id = TRUE)
```

```{r}
#| label: fig-ppPlot
#| fig-cap: "Residual Plots After Transformation of Fantasy Points."
#| fig-alt: "Residual Plots After Transformation of Fantasy Points."

petersenlab::ppPlot(
  linearRegressionModel_outcomeTransformed)
```

@@

Then, what are the regression coefficients?
We do not know what are the correct regression coefficients because each of the possibilities fits the data equally well.
Thus, when estimating the regression model, we could obtain arbitrary estimates of the regression coefficients with an enormous standard error around each estimate.
In general, multicollinearity increases the uncertainty (i.e., standard errors and confidence intervals) around the parameter estimates.
Any [predictor variables](#sec-correlationalStudy) that have a correlation above ~ $r = .30$ with each other could have an impact on the confidence interval of the regression coefficient.
As the correlations among the [predictor variables](#sec-correlationalStudy) increase, the chance of getting an arbitrary answer increases, sometimes called "bouncing betas."
So, it is important to examine a correlation matrix of the [predictor variables](#sec-correlationalStudy) before putting them in the same regression model.
You can also examine indices such as variance inflation factor (VIF), where a value greater than 5 or 10 indicates multicollinearity.

Here is the VIF from our earlier model:

```{r}
car::vif(linearRegressionModel_outcomeTransformed)
```

@@

```{r}
player_stats_seasonal$target_share_log <- log(player_stats_seasonal$target_share + 1)

modelData <- player_stats_seasonal %>% 
  filter(position %in% c("WR")) %>% 
  select(fantasyPoints_transformed, age, height, weight, target_share_log)

numObs <- sum(complete.cases(modelData))
varMeans <- colMeans(modelData, na.rm = TRUE)
varCovariances <- cov(modelData, use = "pairwise.complete.obs")

pairwiseRegression_syntax <- '
  fantasyPoints_transformed ~ age + height + weight + target_share_log
  fantasyPoints_transformed ~~ fantasyPoints_transformed
  fantasyPoints_transformed ~ 1
'

pairwiseRegression_fit <- lavaan(
  pairwiseRegression_syntax,
  sample.mean = varMeans,
  sample.cov = varCovariances,
  sample.nobs = numObs
)

summary(
  pairwiseRegression_fit,
  standardized = TRUE,
  rsquare = TRUE)
```

@@

```{r}
numImputations <- 5

dataToImpute <- player_stats_seasonal %>% 
  filter(position %in% c("WR")) %>% 
  select(player_id, where(is.numeric)) %>% 
  select(
    player_id:games, carries:wopr, fantasy_points, fantasy_points_ppr,
    rush_40_yds, rec_40_yds, fumbles, two_pts, return_yds,
    rush_100_yds:draftround, height:target_share_log) %>% 
  select(-c(fantasy_points_ppr, ageCentered20, target_share_log)) # drop collinear variables

predictors <- c("targets","receiving_yards","receiving_air_yards","receiving_yards_after_catch","receiving_first_downs","racr")

dataToImpute$player_id_integer <- as.integer(as.factor(dataToImpute$player_id))

varsToImpute <- c("age","height","weight","target_share")
Y <- varsToImpute
```

Now, let's specify the imputation method:

```{r}
meth <- make.method(dataToImpute)
meth[1:length(meth)] <- ""
meth[Y] <- "2l.pmm" # specify the imputation method here; this can differ by outcome variable
```

Now, let's specify the prediction matrix.
A predictor matrix is a matrix of values, where:

- columns with non-zero values are predictors of the variable specified in the given row
- the diagonal of the predictor matrix should be zero because a variable cannot predict itself

The values are:

- NOT a predictor of the outcome: `0`
- cluster variable: `-2`
- fixed effect of predictor: `1`
- fixed effect and random effect of predictor: `2`
- include cluster mean of predictor in addition to fixed effect of predictor: `3`
- include cluster mean of predictor in addition to fixed effect and random effect of predictor: `4`

```{r}
pred <- make.predictorMatrix(dataToImpute)
pred[1:nrow(pred), 1:ncol(pred)] <- 0
pred[Y, "player_id_integer"] <- (-2) # cluster variable
pred[Y, predictors] <- 1 # fixed effect predictors
pred[Y, "age"] <- 2 # random effect predictor
pred[Y, Y] <- 1 # fixed effect predictor

diag(pred) <- 0
pred
```

Now, let's run the imputation:

```{r}
imp <- mice(
  as.data.frame(dataToImpute),
  method = meth,
  predictorMatrix = pred,
  m = numImputations,
  maxit = 5, # generally use 100 maximum iterations; this example uses 5 for speed
  seed = 52242)
```

Below are some imputation diagnostics.
Trace plots are in @fig-miTracePlots.

```{r}
#| label: fig-miTracePlots
#| fig-cap: "Trace plots from multiple imputation."
#| fig-alt: "Trace plots from multiple imputation."

plot(imp, c("target_share"))
```

A density plot is in @fig-miDensityPlot.

```{r}
#| label: fig-miDensityPlot
#| fig-cap: "Density plot from multiple imputation."
#| fig-alt: "Density plot from multiple imputation."

densityplot(imp, ~ target_share)
```

The imputated data does not match well the distribution of the observed data.
Thus, it may be necessary to select a different imputation method for more accurate imputation.

Now, let's do some post-processing:

```{r}
imp_long <- complete(
  imp,
  action = "long",
  include = TRUE)

imp_long$target_share_log <- log(imp_long$target_share + 1)

imp_long$fantasyPoints_transformed <- predict(
  yjTransformed,
  newdata = imp_long["fantasyPoints"])$fantasyPoints

imp_mids <- as.mids(imp_long)
```

Now let's estimate multiple regression with the multiply imputed data:

```{r}
imp_regression <- with(
  imp_mids,
  lm(
    fantasyPoints_transformed ~ age + height + weight + target_share_log)
  )

summary(pool(imp_regression))
```

@@

```{r}
fimlRegression_syntax <- '
  fantasyPoints_transformed ~ age + height + weight + target_share_log
  fantasyPoints_transformed ~~ fantasyPoints_transformed
  fantasyPoints_transformed ~ 1
'

fimlRegression_fit <- sem(
  fimlRegression_syntax,
  data = player_stats_seasonal,
  missing = "ML",
  fixed.x = FALSE
)

summary(
  fimlRegression_fit,
  standardized = TRUE,
  rsquare = TRUE)
```

@@

rms::robcov(rms::ols(
  fantasyPoints_transformed ~ age + height + weight + target_share_log,
  data = player_stats_seasonal_subset,
  x = TRUE,
  y = TRUE),
  cluster = player_stats_seasonal_subset$player_id) #account for nested data within player
```

@@

```{r}
bayesianMultipleRegressionModel <- brm(
  formula = fantasyPoints_transformed ~ age + height + weight + I(log(target_share + 1)),
  data = player_stats_seasonal %>% filter(position %in% c("WR")),
  family = gaussian()
)

summary(bayesianMultipleRegressionModel)
```
